---
title: "Monthly Projections"
date: "`r format(Sys.time(), '%d %B, %Y')`"
editor_options:
  chunk_output_type: console
output: html_document
params:
  debug:
    input: bool
    label: Debug Mode
    value: true
  prng_seed:
    input: number
    label: Random Number Seed
    value: 2711
  date_of_valuation:
    input: date
    label: Date of Valuation
    value: "2024-09-30"
  reporting_path:
    input: file
    label: Working directory for the monthly valuation process
    value: ""
  database_name:
    input: file
    label: Database file containing monthly data
    value: ""
  treaty_positions:
    input: file
    label: "Treaty Positions CSV file"
    value: ""
  development_factors:
    input: file
    label: "Development Factors CSV file"
    value: ""
  exceptions_table:
    input: file
    label: "Exceptions Table CSV file"
    value: ""
  key_map:
    input: file
    label: "Key Translations CSV file"
    value: ""
  source_data_table_name:
    input: text
    label: "Name of the source data table"
    value: ""
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = params$debug)
knitr::opts_knit$set(progress = TRUE, verbose = TRUE)
print(params$debug)
```

## Monthly Projections

### Initialize Libraries and Environment

```{r, Initialize Environment}
# Check if libraries are installed and load them
pacman::p_load(
  tidyverse,
  lubridate,
  RSQLite
)

proj.env <- new.env()
set.seed(params$prng_seed) # seed the random number generator
# prevent scientific notation for cross-platform compatibility
options(scipen = 999)

proj.env$time_begin <- Sys.time()
```

### Assumptions

Treaty periods are 12 months (e.g. in WrittenPattern function).

(However, policy length values are defined per treaty in "TREATIES\$Policy.Length".)

Every policy is effective on the first of the month.

LOD Policies

-   Underlying policies are paid at inception, generating UEPR (not monthly) (Inherited UEPR)

-   Does not cede back UEPR at the end of the policy

### Functions

These are more general functions that can be applied throughout the document. In later sections, context specific functions are defined at the beginning of their respective code blocks.

```{r, Localized Functions}
# Return the numeric value of second argument if available.
# Otherwise, return the value of first argument.
# (Used to revise projected values with the reported amounts.)
datareplace <- function(fallback, new_data) {
  if (is.na(new_data)) {
    as.numeric(fallback)
  } else {
    as.numeric(new_data)
  }
}

# Return the number of months between two dates
months_between <- function(EFF, CAL) {
  num_months <- (year(CAL) - year(EFF)) * 12 + (month(CAL) - month(EFF))
  return(num_months)
}

# Convert Excel serial date to date format
convert_excel_date <- function(excel_date) {
  as_date(excel_date, origin = "1899-12-30")
}

# Function to preserve NAs while summing
sum_with_na <- function(x) {
  if (all(is.na(x))) {
    return(NA)
  } else {
    return(sum(x, na.rm = TRUE))
  }
}
```

### Load Data

TREATIES contains many details regarding each treaty (e.g. Effective Date, Prior Loss Ratio, Target Participation, etc...).

DEVS are the development factors used for each treaty.

exc_table contains confirmed revisions to the monthly submission data recorded by our data provider.

```{r, Load Data}
# Import data and ensure strings aren't recognized as factors
proj.env$treaty_positions <- read.csv(
  paste(params$reporting_path, params$treaty_positions, sep = ""),
  stringsAsFactors = FALSE
)
proj.env$development_factors <- read.csv(
  paste(params$reporting_path, params$development_factors, sep = ""),
  stringsAsFactors = FALSE
)
proj.env$exc_table <- read.csv(
  paste(params$reporting_path, params$exceptions_table, sep = ""),
  stringsAsFactors = FALSE
)
```

### Load latest query of data and merge with the Exception Table

```{r, Query Most Recent Experience Data}
proj.env$con <- dbConnect(
  SQLite(),
  dbname = paste(params$reporting_path, params$database_name, sep = "")
)
proj.env$our_db <- dbReadTable(proj.env$con, params$source_data_table_name)
dbDisconnect(proj.env$con)

# Find the latest query_date.
# Store it as a string and filter to only retain matching rows
proj.env$latest_query_date <- format(max(ymd_hm(proj.env$our_db$query_date)), "%Y-%m-%d %H:%M")
proj.env$our_db_filtered <- proj.env$our_db %>% dplyr::filter(query_date == proj.env$latest_query_date)
```

```{r, Translate Source Data to our Index Key}
# Read in translations from program names to program used by our organization
proj.env$key_translations <- read.csv(
  paste(params$reporting_path, params$key_map, sep = ""),
  stringsAsFactors = FALSE
)
```

```{r, Join and Consolidate Data using Keys}
# Join the keys to the dataframe
proj.env$vel_df <- proj.env$our_db_filtered %>%
  left_join(proj.env$key_translations, by = c("nre_feed", "treaty_year"))
```

This step consolidates treaties that were broken out by our data provider into a single treaty.

```{r, Summarize Data by Keys}
# Rows are grouped by KEY and month.
# All numeric values are summed and NAs are preserved.
proj.env$vel_df1 <- proj.env$vel_df %>%
  group_by(KEY, bordereaux_month) %>%
  summarise(across(where(is.numeric), sum_with_na),
    across(where(~ is.numeric(.x)), first),
    .groups = "drop"
  )
```

Rename the columns and subtract salvage & subrogation from Paid Loss

```{r, Adjust the data for experience to date}
proj.env$vel_df2 <- proj.env$vel_df1 %>%
  rename(
    DATAWPARTMONTH = tf_assumed_gross_written_premium,
    DATAEPARTMONTH = tf_assumed_earned_premium,
    DATAPARTPAIDALAEMONTH = tf_assumed_paid_alae,
    DATAPARTCASEOSLOSS = tf_assumed_expected_losses
  ) %>%
  mutate(
    DATAPARTPAIDLOSSMONTH = tf_assumed_paid_losses - coalesce(tf_assumed_ss, 0),
    DATAPARTPAIDULAEMONTH = NA,
    DATAPARTCASEOSALAE = NA,
    DATAPARTCASEOSULAE = NA,
    MONTH = year(bordereaux_month) * 100 + month(bordereaux_month)
  )
```

Consolidate to only keeping relevant columns

```{r, Consolidate Data}
proj.env$vel_df3 <- proj.env$vel_df2 %>%
  select(
    KEY,
    MONTH,
    DATAWPARTMONTH,
    DATAEPARTMONTH,
    DATAPARTPAIDLOSSMONTH,
    DATAPARTPAIDALAEMONTH,
    DATAPARTPAIDULAEMONTH,
    DATAPARTCASEOSLOSS,
    DATAPARTCASEOSALAE,
    DATAPARTCASEOSULAE
  )
```

### Prepare an unabridged Exception Table

The Exception Table only contains a small amount of rows containing data to add or replace to the import from source data provider.

We will expand the Exception Table to contain the full 612 rows for every treaty included in this valuation (a row for each month from 2020 - 2070). These steps will simply retain all original Exception Table rows and add any missing rows with NA values.

```{r, Account for Data Exceptions}
# Create a table with all 612 months for every treaty
proj.env$dates <- seq.Date(as.Date("2020-01-01"), as.Date("2070-12-01"), by = "month")
proj.env$date_vector <- year(proj.env$dates) * 100 + month(proj.env$dates)
proj.env$full_months <- expand.grid(KEY = as.character(proj.env$treaty_positions$KEY.), MONTH = proj.env$date_vector) %>%
  mutate(KEY = as.character(KEY)) %>%
  arrange(KEY)
```

```{r, Not sure}
proj.env$monthly_df <- left_join(proj.env$exc_table, proj.env$full_months, by = c("KEY", "MONTH"))

proj.env$monthly_df <- bind_rows(proj.env$exc_table, proj.env$full_months, .id = "source_id") %>%
  arrange(source_id) %>%
  group_by(KEY, MONTH) %>%
  slice(1) %>%
  ungroup() %>%
  select(-source_id)
```

### Merge source data with expanded table

This step will coalesce our expanded monthly table with the source data. We prioritize values in our exception table where they exist, and resort to source data for all entries in the exception table that are NA.

(Note that a value of 0 is not equivalent to NA. NA indicates that there is no observed data for that value, whereas 0 indicates an observed value of 0.)

```{r, Merge Source Data with Expanded Table}
proj.env$DATATODATE <- proj.env$monthly_df %>%
  full_join(proj.env$vel_df3, by = c("KEY", "MONTH")) %>%
  mutate(
    DATAWPARTMONTH = coalesce(DATAWPARTMONTH.x, DATAWPARTMONTH.y),
    DATAEPARTMONTH = coalesce(DATAEPARTMONTH.x, DATAEPARTMONTH.y),
    DATAPARTPAIDLOSSMONTH = coalesce(DATAPARTPAIDLOSSMONTH.x, DATAPARTPAIDLOSSMONTH.y),
    DATAPARTPAIDALAEMONTH = coalesce(DATAPARTPAIDALAEMONTH.x, DATAPARTPAIDALAEMONTH.y),
    DATAPARTPAIDULAEMONTH = coalesce(DATAPARTPAIDULAEMONTH.x, DATAPARTPAIDULAEMONTH.y),
    DATAPARTCASEOSLOSS = coalesce(DATAPARTCASEOSLOSS.x, DATAPARTCASEOSLOSS.y),
    DATAPARTCASEOSALAE = coalesce(DATAPARTCASEOSALAE.x, DATAPARTCASEOSALAE.y),
    DATAPARTCASEOSULAE = coalesce(DATAPARTCASEOSULAE.x, DATAPARTCASEOSULAE.y)
  ) %>%
  select(KEY, MONTH, DATAWPARTMONTH, DATAEPARTMONTH, DATAPARTPAIDLOSSMONTH, DATAPARTPAIDALAEMONTH, DATAPARTPAIDULAEMONTH, DATAPARTCASEOSLOSS, DATAPARTCASEOSALAE, DATAPARTCASEOSULAE)
```

### Join and Prepare Data

This step creates a dataframe with every treaty having monthly rows from 2020 - 2070.

We also join the premium and loss data from DATATODATE, rename columns, and sort by treaty and month.

```{r, Prepare data}
# Create a vector with monthly dates from 2020 through 2070
proj.env$date_vector <- seq.Date(as.Date("2020-01-01"), as.Date("2070-12-01"), by = "month")

# expand.grid works like a cross_join, but doesn't require the vectors to be dataframes
proj.env$MonthlyProjections <- expand.grid(Treaty = proj.env$treaty_positions$KEY., Month = proj.env$date_vector) %>%
  left_join(proj.env$treaty_positions, by = c("Treaty" = "KEY.")) %>%
  left_join((proj.env$DATATODATE %>% mutate(MONTH = ym(MONTH))), by = c("Treaty" = "KEY", "Month" = "MONTH")) %>%
  rename(
    Program_Name = MGA...Program.Name.,
    Carrier = Carrier.,
    Primary_LOB = Primary.Line.s..of.Business.,
    Secondary_LOB = Secondary.Line.s..of.Business.,
    Effective_Date = Effective.Date.,
    Suffix = Suffix.,
    RA_or_LOD = RA.or.LOD.,
    Total_Subject_Premium = Total.Subject.Premium.,
    Target_Participation = Target.Participation...,
    Inherited_UEPR = Inherited.UEPR...outside.of.participation..,
    ALAE_ULAE_In_Ceded_LR = ALAE..ULAE.or.both.included.in.ceded.loss.ratio.,
    Est_Payout_Duration_Yrs = Estimated.Payout.Duration..years.from.point.exposure..,
    Prior_No_Impr_Gross_LALAE_Ratio_Excl_Corr = Prior.No.Improvement.Gross.LALAE.Ratio..excluding.corridor..LR.cap..etc...,
    Prior_Half_Impr_Gross_LALAE_Ratio_Excl_Corr = Prior.Half.Improvement.Gross.LALAE.Ratio..excluding.corridor..LR.cap..etc...,
    ULAE_Perc = ULAE.Outside.Ceding.Commission...,
    Reins_Broker_Perc = Reinsurance.Broker.Commission...,
    Expenses_Perc = Other.Separately.Charged..Expenses...,
    LALAE_Ratio_No_Improv = Prior.No.Improvment.Net.LALAE.Ratio..including.corridor..LR.cap..etc...,
    LALAE_Ratio_Half_Improv = Prior.Half.Improvment.Net.LALAE.Ratio..including.corridor..LR.cap..etc...,
    LALAE_Ratio_Break_Even = Prior.Breakeven.LALAE.Ratio.,
    Policy_Length = Assumed.Policy.Length.Months.,
    WP_Monthly_Data = DATAWPARTMONTH,
    EP_Monthly_Data = DATAEPARTMONTH,
    Loss_Paid_Monthly_Data = DATAPARTPAIDLOSSMONTH,
    ALAE_Paid_Monthly_Data = DATAPARTPAIDALAEMONTH,
    ULAE_Paid_Monthly_Data = DATAPARTPAIDULAEMONTH,
    Loss_Case_Reserve_Data = DATAPARTCASEOSLOSS,
    ALAE_Case_Reserve_Data = DATAPARTCASEOSALAE,
    ULAE_Case_Reserve_Data = DATAPARTCASEOSULAE
  ) %>%
  mutate(Effective_Date = convert_excel_date(Effective_Date)) %>%
  arrange(Treaty, Month)
```

### Project Monthly Earned Pattern to Inherited UEPR

This step projects a monthly earned pattern to the Inherited UEPR, resulting in the IUEPR_Month column.

This only applies to LOD treaties.

```{r, Project Monthly Earned Pattern to Inherited UEPR}
# Define a function used for projecting Inherited UEPR
# NL: For LOD policies only
#     LOD policies are comprised of two components,
#         1) New Risks, and
#         2) Inherited LOD from prior UW period
# Assumptions:
#     1) all inherited UEPR is on policies paid at policy inception
#     2) The average underlying policy is written mid month
# EFF: the effective date of the treaty (assumed to be the first of the month)
# PLEN: Policy Length in Months -- the assumed length of the underlying exposure
# K: the difference between the month in question and the effective date of the treaty
# When this function is used below, it uses column "Policy_Length" for that treaty's assumed policy length
Inherited_UEPR_Pattern <- function(EFF, CAL, PLEN) {
  k <- months_between(EFF, CAL)
  if (k >= 0 & k <= (PLEN - 1))
  # {(12-(k+0.5)*(12/PLEN))/(6*PLEN)}
    {
      (2 / PLEN) * (1 - (k + 0.5) / PLEN)
    }
  # k+0.5 adjusts for the mid-month bind
  # (k+0.5)/PLEN yields the proportion of the overall length that has already passed (Parallelogram horizontal distance)
  # 1-(k+0.5)/PLEN yields the portion of the overall length yet to pass (Parallelogram horizontal distance)
  else {
    0
  }
}

# Project IUEPR by month using Inherited_UEPR_Pattern function
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  rowwise() %>%
  mutate(
    IUEPR_Total = replace_na(Inherited_UEPR, 0),
    IUEPR_Perc = Inherited_UEPR_Pattern(Effective_Date, Month, Policy_Length),
    IUEPR_Month = IUEPR_Total * IUEPR_Perc
  ) %>%
  ungroup()
```

The following is a check to ensure that LOD Treaties have their sum of IUEPR_Month equal to IUEPR_Total.

```{r, Validate Loss Occurred During Treaty Period}
proj.env$MonthlyProjections %>%
  distinct(Treaty, .keep_all = TRUE) %>%
  summarize(Total_IUEPR = sum(IUEPR_Total)) %>%
  pull(Total_IUEPR)

sum(proj.env$MonthlyProjections$IUEPR_Month)
```

### Project Company's Share of Monthly Written Premium

This step projects share of monthly written premium and revises projections with reported data.

The projection evenly distributes the estimated total written premium across the 12 month treaty period.

```{r, Project Business Proportion of Written Premium}
# Define a function used for projecting monthly written premium
# Calculates the expected fraction of written premium in a given month by checking if that month
# is between the effective date and the expiration date (assuming a 12 month Treaty period)
# EFF = The effective date of the policy, assumed to be the first of the month
# CAL = The calendar month for which we are calculating the expected written premium, assumed to be the first of the month
WrittenPattern <- function(EFF, CAL) {
  treaty_month <- months_between(EFF, CAL)
  if (treaty_month >= 0 & treaty_month <= 11) {
    1 / 12
  } else {
    0
  }
}

# Project Written Participation by Month
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  rowwise() %>%
  dplyr::mutate(
    # Our QS (quota share) of the Estimated Total Written Premium
    WP_Total = Total_Subject_Premium * Target_Participation,
    # Allocate 1/12 for all months during the treaty period, 0 otherwise
    WP_Perc = WrittenPattern(Effective_Date, Month),
    # Project monthly written participation
    WP_Monthly = WP_Total * WP_Perc,
  ) %>%
  ungroup()

# Record the total difference between projected written premiums and those that were reported
# (this isolated step can be consolidated if there's no problem with adding an extra WP_Proj
# column to retain the original projected values)
proj.env$WP_Proj_vs_Rept <- proj.env$MonthlyProjections %>%
  dplyr::filter(!is.na(WP_Monthly_Data)) %>%
  mutate(WP_Proj_vs_Rept = WP_Monthly - WP_Monthly_Data) %>%
  summarize(WP_Proj_vs_Rept_Total = sum(WP_Proj_vs_Rept)) %>%
  pull(WP_Proj_vs_Rept_Total)

# Replace projected data with actual data where available
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  rowwise() %>%
  mutate(WP_Monthly = datareplace(WP_Monthly, WP_Monthly_Data)) %>%
  ungroup()
```

The following is a check to ensure the originally anticipated WP participation (WP_Total) carried forward through the calculations.

```{r, Validate Written Premium Participation}
# Sum of anticipated treaty WP participations
proj.env$MonthlyProjections %>%
  distinct(Treaty, .keep_all = TRUE) %>%
  summarize(Total_WP_Participation = sum(WP_Total)) %>%
  pull(Total_WP_Participation)

# WP_Monthly contains projected monthly WP and, where available, reported WP
# Therefore, we sum the revised WP_Monthly column and add the monthly discrepancies between
# projected and actual reported values.
sum(proj.env$MonthlyProjections$WP_Monthly) + proj.env$WP_Proj_vs_Rept
```

### Project Company's Share of Monthly Earned Premium

This step projects share of monthly earned premium and revises projections with reported data. The calculation for monthly earned premiums uses the updated WP_Monthly from the previous section.

Two temporary tables are created to form and calculate the sumproduct of earned premium for each month, which are then joined back to MonthlyProjections.

```{r, Project Business Proportion of Earned Premium}
# Define a function used for used for projecting monthly earned premium
# k: month of treaty
# PLEN: Policy Length in Months - the assumed length of the underlying exposure (passed from "Policy_Length")
# The function generates a vector with proportions evenly distributed across the assumed months
# of policy length. The 1st month, however, splits half of its share with the month following the policy duration.
EarnedPattern <- function(k, PLEN) {
  case_when(
    (k >= 1 & k <= (PLEN - 1)) ~ 1 / PLEN,
    (k == 0 | k == PLEN) ~ 1 / (2 * PLEN),
    TRUE ~ 0
  )
}

# Create first table containing all rows in MonthlyProjections dataframe
proj.env$table1_Earning_Pattern <- proj.env$MonthlyProjections %>%
  select(
    # Keys
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    WP_Month = Month,
    # Attribute
    Policy_Length,
    WP_Monthly
  )

# Create a second table by joining on the newly created EP_Month, resulting in multiple earning months for each WP month
proj.env$table2_Earning_Pattern <- left_join(
  proj.env$table1_Earning_Pattern,
  select(proj.env$table1_Earning_Pattern, EP_Month = WP_Month) %>% distinct(),
  join_by(WP_Month <= EP_Month)
) %>%
  # Calculate the lag between writing and earning
  dplyr::mutate(Earned_Lag = months_between(WP_Month, EP_Month)) %>%
  # Calculate the percent of premium earned in each month from each WP month
  dplyr::mutate(Earned_Pct = EarnedPattern(Earned_Lag, Policy_Length)) %>%
  # Earned Premium from each WP month is the product of that month's WP
  dplyr::mutate(EP_Monthly = Earned_Pct * WP_Monthly) %>%
  # Remove all rows that don't have an Earned Premium. (And retain negative earned premiums.)
  dplyr::filter(EP_Monthly != 0) %>%
  # Sum across WP months because it's irrelevant which WP month it was sourced from, rather only when it was earned
  group_by(
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    EP_Month
  ) %>%
  summarize(EP_Monthly = sum(EP_Monthly)) %>%
  ungroup()

# Join the earned premiums (EP_Monthly) to MonthlyProjections
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  left_join(select(proj.env$table2_Earning_Pattern, c(Program_Name, Primary_LOB, Effective_Date, Suffix, EP_Month, EP_Monthly)),
    by = c(
      "Program_Name",
      "Primary_LOB",
      "Effective_Date",
      "Suffix",
      "Month" = "EP_Month"
    )
  ) %>%
  mutate(EP_Monthly = replace_na(EP_Monthly, 0))


proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>%
  mutate(
    # Projected Earned Participation by Month = Inherited UEPR + Exposure Policy Earned Premium
    EP_Monthly_Incl_IUEPR = IUEPR_Month + EP_Monthly,
    # Replace Estimated Earned Premium with Reported Earned Premium
    EP_Monthly_Incl_IUEPR = mapply(datareplace, EP_Monthly_Incl_IUEPR, EP_Monthly_Data),
  ) %>%
  group_by(Treaty) %>%
  mutate(
    # Total Earned Premium To Date (for each treaty)
    EP_To_Date = cumsum(EP_Monthly_Incl_IUEPR),
    # Total Written Premium To Date (for each treaty)
    WP_To_Date = cumsum(WP_Monthly)
  ) %>%
  ungroup() %>%
  # UEPR = [Total Inherited UEPR] + [Total Written Premium to date] - [Total Earned Premium to date]
  # [NOTE] 4 of the treaties have negative UEPR values ranging from -50,000 to -500,000 once you
  #        go 1-2 years past treaty expiration
  mutate(UEPR = IUEPR_Total + WP_To_Date - EP_To_Date)
```

```{r, Validate Written to Earned Premium}
# Check that total earned premium == total written premium
sum(proj.env$MonthlyProjections$WP_Monthly)
sum(proj.env$MonthlyProjections$EP_Monthly)
```

### Undeveloping Earned Premiums for Projecting Paid and Reported Losses

This step projects paid and reported losses by undeveloping the total earned premium (including Inherited UEPR where relevant).

Two temporary tables are used to form and calculate the sumproducts of the losses that occur for each month, which are then joined back to MonthlyProjections.

```{r, Apply development factors}
# Prepare the development factors dataset to have 0 months lag indicating the current month
proj.env$DEVS_rev <- proj.env$development_factors %>%
  arrange(KEY, LAG_ACCMONTH) %>%
  dplyr::filter(LAG_ACCMONTH != 0) %>% # Remove rows where LAG_ACCMONTH is 0
  mutate(LAG_ACCMONTH = LAG_ACCMONTH - 1) # Decrement LAG_ACCMONTH by 1

# Create an initial table that will be used for calculating losses
proj.env$t_Loss <- proj.env$MonthlyProjections %>%
  select(
    # Primary Key
    Treaty,
    # Unique Key
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    Effective_Date,
    Month,
    # Attribute
    EP_Monthly_Incl_IUEPR
  ) %>%
  group_by(Treaty) %>%
  mutate(Total_Earned_Partic = sum(EP_Monthly_Incl_IUEPR)) %>%
  ungroup()

# Join the development factors
proj.env$t_Loss_2 <- proj.env$t_Loss %>%
  full_join(proj.env$DEVS_rev, by = c("Treaty" = "KEY"), relationship = "many-to-many") %>%
  dplyr::mutate(
    # This column contains the month for the paid and reported losses
    LAG_date = Month %m+% months(LAG_ACCMONTH),
    # Paid and Reported losses using development factors on Earned Premium
    # I renamed Paid_Loss -> EP_Undev_Paid
    # I renamed Reported_Loss -> EP_Undev_Reported
    EP_Undev_Paid = EP_Monthly_Incl_IUEPR * PAID_PERCENT,
    EP_Undev_Reported = EP_Monthly_Incl_IUEPR * REPORTED_PERCENT
  ) %>%
  # Remove the rows that are past our data's time frame
  dplyr::filter(LAG_date <= as.Date("2070-12-01")) %>%
  # Sum across loss months because it's irrelevant which EP month it was sourced from,
  # rather only when the loss is paid/reported
  group_by(
    Program_Name,
    Primary_LOB,
    Effective_Date,
    Suffix,
    Total_Earned_Partic,
    LAG_date
  ) %>%
  summarise(
    EP_Undev_Paid = sum(EP_Undev_Paid),
    EP_Undev_Reported = sum(EP_Undev_Reported)
  ) %>%
  ungroup() %>%
  # Ensure all months past the scope of the development factors have the full paid and reported amounts
  mutate(
    EP_Undev_Paid = ifelse(months_between(Effective_Date, LAG_date) >= 240,
      Total_Earned_Partic, EP_Undev_Paid
    ),
    EP_Undev_Reported = ifelse(months_between(Effective_Date, LAG_date) >= 240,
      Total_Earned_Partic, EP_Undev_Reported
    )
  ) %>%
  ungroup() %>%
  # Remove columns that won't be imported
  select(-Total_Earned_Partic)

proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>% left_join(proj.env$t_Loss_2, by = c(
  "Program_Name",
  "Primary_LOB",
  "Effective_Date",
  "Suffix",
  "Month" = "LAG_date"
))
```

The following is a check to ensure the columns for paid and reported losses sum to the total undeveloped earned premium. There may be slight deviances due to the different calculations involved with each field.

```{r, Validate Losses}
proj.env$Check_EP_Undev_Losses <- proj.env$MonthlyProjections %>%
  group_by(Treaty) %>%
  summarize(
    EP_Monthly_Incl_IUEPR_Total = sum(EP_Monthly_Incl_IUEPR),
    EP_Undev_Paid_Total = max(EP_Undev_Paid),
    EP_Undev_Reported_Total = max(EP_Undev_Reported)
  )

sum(proj.env$Check_EP_Undev_Losses$EP_Monthly_Incl_IUEPR_Total)
sum(proj.env$Check_EP_Undev_Losses$EP_Undev_Paid_Total)
sum(proj.env$Check_EP_Undev_Losses$EP_Undev_Reported_Total)
```

### Create Additional Projections

This step calculates the dollar amounts for ULAE, reinsurance broker commission, expenses, and multiple LALAE values.

```{r, Additional Projections}
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>% mutate(
  ULAE_Total = EP_To_Date * ULAE_Perc,
  Reins_Broker_Comm_Total = EP_To_Date * Reins_Broker_Perc,
  Expenses_Total = EP_To_Date * Expenses_Perc,
  LALAE_Paid_Total_No_Improv = EP_Undev_Paid * LALAE_Ratio_No_Improv,
  LALAE_Paid_Total_Half_Improv = EP_Undev_Paid * LALAE_Ratio_Half_Improv,
  LALAE_Paid_Total_Break_Even = EP_Undev_Paid * LALAE_Ratio_Break_Even,
  LALAE_Reported_Total_No_Improv = EP_Undev_Reported * LALAE_Ratio_No_Improv,
  LALAE_Reported_Total_Half_Improv = EP_Undev_Reported * LALAE_Ratio_Half_Improv,
  LALAE_Reported_Total_Break_Even = EP_Undev_Reported * LALAE_Ratio_Break_Even,
  LALAE_Reserve_No_Improv = LALAE_Ratio_No_Improv * (EP_To_Date - EP_Undev_Paid),
  LALAE_Reserve_Half_Improv = LALAE_Ratio_Half_Improv * (EP_To_Date - EP_Undev_Paid),
  LALAE_Reserve_Break_Even = LALAE_Ratio_Break_Even * (EP_To_Date - EP_Undev_Paid),
  LALAE_IBNR_No_Improv = LALAE_Ratio_No_Improv * (EP_To_Date - EP_Undev_Reported),
  LALAE_IBNR_Half_Improv = LALAE_Ratio_Half_Improv * (EP_To_Date - EP_Undev_Reported),
  LALAE_IBNR_Break_Even = LALAE_Ratio_Break_Even * (EP_To_Date - EP_Undev_Reported)
)
```

To accommodate the Valuation Template, the Month and Effective_Date columns are converted back to YYYYMM format.

```{r, Convert dates to strings}
# Convert Month and Effective_Date to numeric yyyymm format for export
proj.env$MonthlyProjections <- proj.env$MonthlyProjections %>% mutate(
  Month = year(Month) * 100 + month(Month),
  Effective_Date = year(Effective_Date) * 100 + month(Effective_Date)
)
```

### Export Monthly Projections csv

```{r, Export Monthly Projections}
# Export the Monthly Projections to a csv file

# Get today's date in YYYYMMDD format
proj.env$today <- format(Sys.Date(), "%Y%m%d")

# Create the file name
proj.env$file_name <- paste0(
  params$reporting_path,
  proj.env$today, "_",
  format(as.Date(params$date_of_valuation, "%Y-%m-%d"), "%Y%m%d"),
  "_VAL_MONTHLY_PROJECTIONS.csv"
)

# Export the Monthly Projections to a csv file with the dynamic file name
write.csv(
  proj.env$MonthlyProjections,
  file = proj.env$file_name,
  row.names = FALSE
)
```
